---
title: "Tidy Tuesday: 2022, Week 8, Freedom in the World"
author: "Fiona Lees"
date: 2022-02-22
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      out.width = "100%",
                      fig.align = "center")

```

## Freedom in the World

**Note added 1st Mar 2022**: Index such as the "Freedom Index" have potential for bias and/or miscalculation. Data such as these are far from perfect and may misrepresent nuanced political situations or oversimplify difficult to record/measure political nuance.  

The [data](https://freedomhouse.org/report/freedom-world) this week comes from [Freedom House](https://freedomhouse.org/). It provides freedom index scores and freedom statuses for 195 countries and 15 territories. Global freedom statuses are calculated on a weighted scale, based on 10 political rights indicators and 15 civil liberties indicators.The combination of the overall score awarded for political rights and the overall score awarded for civil liberties, after being equally weighted, determines a status of Not Free, Partly Free, or Free. More information on the methodology is provided [here](https://freedomhouse.org/reports/freedom-world/freedom-world-research-methodology).  

[Arthur Cheib](https://github.com/ArthurCheib/analytical-politics-project/blob/main/data/tidy-data-fh-un.csv) kindly made the dataset available to Tidy Tuesday. However, after spending a bit of time reading about the data and how it's collected, I decided to pull a fuller dataset directly from the Freedom House website. This was mainly so I could use the overall index scores in my analysis rather than the separate ratings for civil liberties and political rights.  


### Objectives
A few weeks ago (9th Feb) I saw a [piece in the Economist](https://www.economist.com/graphic-detail/2022/02/09/a-new-low-for-global-democracy) about global democracy. It contained a beeswarm chart I really liked, showing a global democracy index for 167 countries. The challenge I set myself this week was to try to produce a similar chart using Freedom House's freedom index.  


### Learning Points
I learned about a couple of new things this week:  

- **ggbeeswarm package:** I learned how to use geom_beeswarm() to produce a beeswarm chart and position_beeswarm() to position text labels beside points within the chart. It took me a wee while and a bit of messing about to position the text labels the way I wanted, but I got there in the end. I did consider using ggplot2::geom_jitter(), ggforce::geom_sina() and ggdist::stat_dots() to create the beeswarm chart, but in the end I found ggbeeswarm::geom_beeswarm() gave me the look I wanted. There's a really good tutorial on [Cedric Scherer's website](https://www.cedricscherer.com/2021/06/06/visualizing-distributions-with-raincloud-plots-and-how-to-create-them-with-ggplot2/) that covers several of these options.  

- **MetBrewer package** I've been eager to try out Blake R Mills amazing MetBrewer colour palettes ever since they became available. I finally had a good reason this week. After considering all the colour blind friendly palettes, I found [Hiroshige](https://github.com/BlakeRMills/MetBrewer#hiroshige) was a good option for this week's visualisation.  


### Load Packages

```{r packages}

library(showtext)
library(ggbeeswarm)
library(MetBrewer)
library(tidyverse)

```


### Import Data
I've saved the [MS Excel file on Freedom House's website](https://freedomhouse.org/report/freedom-world) to my local hard drive, so I'm going to load it from there. I'll do a little tidying of column names as I import the data.  

```{r import_data_2}

freedom <- readxl::read_xlsx(
  "data/Aggregate_Category_and_Subcategory_Scores_FIW_2003-2021.xlsx",
  range = "FIW06-21!A1:S3340",
  # Explicitly define the data type in each column to ensure it gets pulled in correctly
  col_types = c("text", "text", "text", "numeric", "text", "numeric", "numeric", 
                "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", 
                "numeric", "numeric", "numeric", "numeric", "numeric", "numeric")
  ) %>% 
 janitor::clean_names()

glimpse(freedom)

```

Let's have a quick look at the `freedom` data. The freedom index score I'm interested in is stored in the variable called `total`.      

```{r summary_2}

view(freedom)
Hmisc::describe(freedom)

```


### Explore / Wrangle Data

I'm only going to use the most recent country level data, so I'll filter out the earlier years and data for territories. While I'm at it, I'll drop the variables I'm not interested in at the moment.  

```{r wrangle_1}

# Drop data not required
countries_2020 <- freedom %>% 
  filter(c_t == "c" & edition == 2021) %>% 
  select(country_territory:cl_rating, pr, cl, total) %>% 
  rename(country = country_territory)

```

Next, I'd like to get a sense of the range of scores in each region of the world.  

```{r explore_1}
# How many countries in each region? 
# What are the range of total scores in each region?
countries_2020 %>% 
  group_by(region) %>% 
  summarise(n_countries = n(),
            min_score = min(total),
            max_score = max(total),
            median_score = median(total),
            mean_score = mean(total),
            stdev_score = sd(total)

  ) %>% 
  ungroup() %>% 
  arrange(mean_score)

```

Okay, now I'm going to change `status` and `region` into factors and give each factor level a meaningful (formatted) label. This will make these variables much easier to work with when I create my chart.  

```{r wrangle_2}

# Make status and region factors
countries_2020 <- countries_2020 %>% 
  mutate(year = edition - 1, .after = edition,
         status = factor(status, levels = c("NF", "PF", "F"), 
                         labels = c("Not Free", "Partly Free", "Free")),
         region_2 = factor(region, levels = c("MENA", "Eurasia", "SSA", "Asia", "Americas", "Europe"), 
                         labels = c("Middle East &\nNorth Africa", "Eurasia", 
                                    "Sub-Saharan\nAfrica", "Asia &\nAustralasia", 
                                    "North America,\nLatin America\n& the Caribbean", "Europe")),
         )


 
```

Finally, I want to separate out the top and bottom ranked countries in each region, so that I can label them in my chart.  

```{r wrangle_3}

# Highest score(s) in each region
countries_2020_high <- countries_2020 %>% 
  group_by(region) %>% 
  slice(which(total == max(total))) %>% 
  ungroup() %>% 
  arrange(region_2, total)
  
# Lowest score(s) in each region
countries_2020_low <- countries_2020 %>% 
  group_by(region) %>% 
  slice(which(total == min(total))) %>% 
  ungroup() %>% 
  arrange(region_2, total)

# Create a flag for the highest and lowest countries in the `countries_2020` dataset
countries_2020 <- countries_2020 %>%
  mutate(hi_low_country = ifelse(country %in% countries_2020_high$country | country %in% countries_2020_low$country, TRUE, FALSE))

```


### Visualise Data
I'm ready to start building my visualisation.    

First, I'll load the fonts I want to use.   

```{r fonts}

# Fonts
showtext_auto(enable = TRUE)
font_add_google("Lato")
font_add_google("Roboto Slab")

```

Now I'll create the chart.    

```{r p_final, fig.width = 12, fig.height = 9.5, fig.alt="Beeswarm chart showing global freedom index scores for 195 countries in six world regions (range 0 - 100, 100 being the most free). The freedom status of each country (not free, partly free, free) is also shown. Overall, scores are highest in Europe (Finland, Norway and Sweden = 100), whereas scores in the Middle East and North Africa are the lowest (Syria = 1)."}

p_final <- countries_2020 %>% 
  ggplot(aes(x = total, y = region_2, fill = status, colour = hi_low_country)) +
  # Beeswarm
  geom_beeswarm(
    shape = 21, 
    cex = 1.8, 
    size = 3.8, 
    groupOnX = FALSE
    ) +
  # Dot fill colour and borders
  scale_fill_manual(values = met.brewer("Hiroshige", 3)) +
  scale_colour_manual(values = c("darkgrey", "black"), guide = "none") +
  # Text labels for lowest scores
  geom_label(
    data = countries_2020_low, 
    aes(label = country), 
    colour = "grey40",
    fill = "white",
    size = 3.88, 
    # Use position_beeswarm() to match labels to points
    position = position_beeswarm(groupOnX = FALSE, cex = 1.8),
    # Adjust label positions for individual points
    vjust = c(-0.8, -0.8, 1.5, -0.8, -0.8, 1.5, -0.8), 
    hjust = 1, 
    label.padding = unit(0.02, "cm"),
    label.size = NA
    ) +
  # Text labels for highest scores
  geom_label(
    data = countries_2020_high, 
    aes(label = country), 
    colour = "grey40", 
    fill = "white",
    size = 3.88, 
    # Use position_beeswarm() to match labels to points
    position = position_beeswarm(groupOnX = FALSE, cex = 1.8),
    # Adjust label positions for individual points
    hjust = c(0, 0, 0, 0, 0, 0, 0, -0.25, 0),
    vjust = c(1.8, 2.6, 1.6, 1.6, 1.6, -0.8, -1.9, 1.5, 1.6),
    label.padding = unit(0.02, "cm"),
    label.size = NA
    ) +
  # Horizontal lines between each region  
  geom_hline(yintercept = seq(0.5, 6.5, 1), colour = "grey90") +
  # Scales
  scale_x_continuous(limits = c(-5, 105), breaks = seq(0, 100, 10), position = "top") +
  scale_y_discrete(expand = expansion(add = 0.5)) +
  # Styling
  theme_minimal() +
  theme(
    text = element_text(family = "Lato", colour = "grey40", size = 12),
    plot.title = element_text(family = "Roboto Slab", colour = "black", size = 16, 
                              face = "bold", margin = margin(b = 5), hjust = 0),
    plot.subtitle = element_text(size = 12, margin = margin(b = 0), hjust = 0),
    plot.caption = element_text(size = 10, margin = margin(t = 10), hjust = 0),
    plot.background = element_rect(colour = "white", fill = "white"),
    axis.ticks.y = element_blank(),
    axis.text = element_text(size = 12),
    axis.text.y = element_text(margin = margin(r = 5)),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_line(colour = "grey90"),
    panel.grid.minor.x = element_blank(),
    legend.position = "top",
    legend.justification = "right",
    legend.margin = margin(t = 0, r = 0, b = 0, l = 0),
    legend.box.margin = margin(t = -15, r = 0, b = -15, l = 0),
    legend.text = element_text(size = 12),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10)
    ) +
  # Titles
  labs(
    title = "Global Freedom Index, 2020",
    subtitle = "Freedom index / status of 195 countries, 100 = most free",
    x = "", 
    y = "",
    colour = "",
    fill = "",
    caption = "Note: These type of index have potential for bias / miscalculation. They are far from perfect and may misrepresent / oversimplify nuanced political situations.\n \nGlobal freedom statuses are calculated on a weighted scale, based on 10 political rights indicators and 15 civil liberties indicators.\nThe combination of the overall score awarded for political rights and the overall score awarded for civil liberties, after being equally\nweighted, determines the status of Not Free, Partly Free, or Free. Data based on status Jan 1 - Dec 31 2020.\n \nVisualisation: Fiona Lees (@Fi_Lees) | Source: Freedom House | Tidy Tuesday: Week 8, 2022"
    )

p_final

```

Save the plot as an image.  

```{r save_image}

# Tell showtext to use DPI of 300 so text size is correct when chart is saved as an image
showtext_opts(dpi = 300)

# Save plot as image
ggsave("freedom_2020.png", p_final, width = 12 , height = 9.5, units = "in", dpi = 300)

# Turn off showtext
showtext_auto(FALSE)

```


### Session Information
```{r session_info}

sessionInfo()

```
